---
title       : Data Science Capstone
subtitle    : Next Word Prediction
author      : Pei-Chun Su
job         : Student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---  
  
## Outline  
  
* Prediction Algorithm  
 * Efficient Modelling - Markov Chain / Katz back-off  
 * Cleaned/Compressed Datasets - 580MB -> 36MB  
 * Quick predictive response - 0.000~0.003s  
* Instructions  
 * Input - Sentence (truncate the last 1~4 words)   
 * Output - Top possible words / Wordcloud visualization  
* Experience of Application  
 * User Interface - Shiny server / Amazon EC2  
 * Manual / Documents  

    
--- 

## SELECT THE CAR FOR TRIP APP

This app helps you to choose a car for a trip using the mtcars dataset.
First, you need to inform the distance of your trip and the price of gasoline in your region. These information will be used to calculate the Gasoline Expenditure for each car in the dataset. Then, you can enter the maximum amount of money you want to spend on gasoline, and the table shows only the cars that have Miles per Gallon (mpg) that can go below this value.  
Second, you can choose some caractheristcs of the cars that you desire: Cylinders, Displacement, Horse Power and Transmission. The table will show only the cars with the filters selected. You can sort the table according to the variable you want by clicking the arrows at the top of the table. 

--- 

## MTCARS DATASET

The data used in the app comes from the Motor Trend Car Road Tests (mtcars) dataset. The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). We can look to some carachteristics of the data:  

```{r}
head(mtcars)
```

--- 

## PLOT

Here we can see the relationship between three variables: miles per gallon (mpg), displacement (disp) and weight (wt).
```{r}
library(car)
scatterplotMatrix(~mpg+disp+wt, data=mtcars, cex.axis=1.5)
```
---